---
title: "Difference-in-Differences Analysis"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(fixest)
library(did)
library(dplyr)
library(ggplot2)
library(tibble)
library(purrr)
library(broom)

# set working directory if needed
# setwd("path/to/project")

df <- read_csv("data/processed/for_analysis/va_did_panel.csv")

df <- df %>%
  mutate(
    census_geoid = as.character(census_geoid),
    year = as.integer(year),
    first_entry_year = as.numeric(first_entry_year),
    weight = as.numeric(weight),
    ever_treated = if_else(!is.na(first_entry_year) & first_entry_year > 0, 1, 0),
    post = if_else(ever_treated == 1 & year >= first_entry_year, 1, 0),
    g = if_else(ever_treated == 1, as.integer(first_entry_year), 0L)
  )
```

```{r twfe, message=FALSE, warning=FALSE}
  # Baseline TWFE model
  twfe <- feols(
  inefficient_pricing ~ post | census_geoid + year,
  data = df,
  weights = ~ weight,
  cluster = ~ census_geoid
)

summary(twfe)
```


```{r sun-abraham, message=FALSE, warning=FALSE}
# Sun and Abraham (2021) model

max_year <- max(df$year, na.rm = TRUE)

df <- df %>%
  mutate(
    cohort_sa = if_else(
      ever_treated == 1,
      as.integer(first_entry_year),
      10000L
    ),
    g = if_else(
      ever_treated == 1,
      as.integer(first_entry_year),
      0L
    )
  )

df %>%
  distinct(census_geoid, ever_treated, first_entry_year, cohort_sa, g) %>%
  count(ever_treated, cohort_sa, g) %>%
  arrange(ever_treated, cohort_sa)

sa <- feols(
  inefficient_pricing ~ sunab(cohort_sa, year, ref.p = -1) |
    census_geoid + year,
  data = df,
  weights = ~ weight,
  cluster = ~ census_geoid
)

summary(sa)
summary(sa, agg = "ATT")
iplot(sa)
```

## Callaway and Sant'Anna (2021)

```{r callaway-santanna, message=FALSE, warning=FALSE}
library(did)
library(dplyr)

# ------------------------------------------------------------
# Clean dataframe for did::att_gt()
# ------------------------------------------------------------

df_cs <- df %>%
  mutate(
    id_num = as.numeric(as.integer(factor(census_geoid))),
    year = as.numeric(year),
    g = if_else(ever_treated == 1, as.numeric(first_entry_year), 0),
    weight = as.numeric(weight),
    inefficient_pricing = as.numeric(inefficient_pricing)
  ) %>%
  select(id_num, year, g, weight, inefficient_pricing) %>%
  filter(
    !is.na(id_num),
    !is.na(year),
    !is.na(g),
    !is.na(weight),
    !is.na(inefficient_pricing),
    weight > 0
  ) %>%
  as.data.frame()

# Quick checks
df_cs %>%
  summarise(
    rows = n(),
    ids = n_distinct(id_num),
    years = n_distinct(year),
    treated_ids = n_distinct(id_num[g > 0]),
    control_ids = n_distinct(id_num[g == 0])
  )

df_cs %>%
  distinct(id_num, g) %>%
  count(g)
```

```{r callaway-main, message=FALSE, warning=FALSE}
# ------------------------------------------------------------
# Estimate ATT(g,t)
# ------------------------------------------------------------

cs <- did::att_gt(
  yname = "inefficient_pricing",
  tname = "year",
  idname = "id_num",
  gname = "g",
  data = df_cs,
  panel = TRUE,
  allow_unbalanced_panel = TRUE,
  control_group = "nevertreated",
  xformla = ~ 1,
  est_method = "reg",
  weightsname = "weight",
  bstrap = FALSE,
  faster_mode = FALSE
)

summary(cs)
```

```{r callaway-dynamic, message=FALSE, warning=FALSE}
# ------------------------------------------------------------
# Dynamic / event-study aggregation
# ------------------------------------------------------------

cs_dyn <- did::aggte(
  cs,
  type = "dynamic",
  na.rm = TRUE
)

summary(cs_dyn)

did::ggdid(cs_dyn) +
  ggplot2::labs(
    title = "Callaway-Sant'Anna Dynamic Effects",
    x = "Event time",
    y = "ATT"
  )
```

```{r callaway-simple, message=FALSE, warning=FALSE}
# ------------------------------------------------------------
# Overall simple ATT
# ------------------------------------------------------------

cs_simple <- did::aggte(
  cs,
  type = "simple",
  na.rm = TRUE
)

summary(cs_simple)
```

```{r compare models, message=FALSE, warning=FALSE}
## Compare DiD estimators

```{r did-method-comparison-plot, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(tibble)

# -----------------------------
# Extract estimates
# -----------------------------

# TWFE
twfe_est <- broom::tidy(twfe, conf.int = TRUE) %>%
  filter(term == "post") %>%
  transmute(
    method = "TWFE",
    estimate = estimate,
    se = std.error,
    conf.low = conf.low,
    conf.high = conf.high
  )

# Sun-Abraham aggregate ATT
sa_att <- summary(sa, agg = "ATT")

sa_est <- tibble(
  method = "Sun-Abraham",
  estimate = coef(sa_att)[1],
  se = se(sa_att)[1],
  conf.low = estimate - 1.96 * se,
  conf.high = estimate + 1.96 * se
)

# Callaway-Sant'Anna simple ATT
cs_est <- tibble(
  method = "Callaway-Sant'Anna",
  estimate = cs_simple$overall.att,
  se = cs_simple$overall.se,
  conf.low = estimate - 1.96 * se,
  conf.high = estimate + 1.96 * se
)

# Combine
did_compare <- bind_rows(
  twfe_est,
  sa_est,
  cs_est
) %>%
  mutate(
    method = factor(
      method,
      levels = c("TWFE", "Sun-Abraham", "Callaway-Sant'Anna")
    )
  )

did_compare
```

```{r did-method-comparison-plot, message=FALSE, warning=FALSE}
ggplot(did_compare, aes(x = method, y = estimate)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = conf.low, ymax = conf.high),
    width = 0.15,
    linewidth = 0.8
  ) +
  coord_flip() +
  labs(
    title = "Estimated Effect of Data Center Entry on Inefficient Pricing",
    subtitle = "Comparison across DiD estimators",
    x = NULL,
    y = "Estimated ATT ($/MWh)"
  ) +
  theme_minimal(base_size = 13)
  ```

  ## Multi-outcome DiD models

```{r multi-outcome-did, message=FALSE, warning=FALSE}
# ------------------------------------------------------------
# Outcomes to estimate
# ------------------------------------------------------------

outcomes <- c(
  "avg_total_lmp",
  "inefficient_pricing",
  "avg_congestion",
  "avg_loss",
  "avg_energy"
)

outcome_labels <- c(
  inefficient_pricing = "Inefficient Pricing",
  avg_congestion = "Congestion",
  avg_loss = "Marginal Loss",
  avg_total_lmp = "Total LMP",
  avg_energy = "System Energy"
)

# ------------------------------------------------------------
# Helper: TWFE
# ------------------------------------------------------------

run_twfe <- function(y, data = df) {
  feols(
    as.formula(paste0(y, " ~ post | census_geoid + year")),
    data = data,
    weights = ~ weight,
    cluster = ~ census_geoid
  )
}

extract_twfe <- function(model, y) {
  broom::tidy(model, conf.int = TRUE) %>%
    filter(term == "post") %>%
    transmute(
      outcome = y,
      method = "TWFE",
      estimate = estimate,
      se = std.error,
      conf.low = conf.low,
      conf.high = conf.high,
      p.value = p.value
    )
}

# ------------------------------------------------------------
# Helper: Sun-Abraham
# ------------------------------------------------------------

run_sa <- function(y, data = df) {
  feols(
    as.formula(
      paste0(y, " ~ sunab(cohort_sa, year, ref.p = -1) | census_geoid + year")
    ),
    data = data,
    weights = ~ weight,
    cluster = ~ census_geoid
  )
}

extract_sa <- function(model, y) {
  sa_att <- summary(model, agg = "ATT")
  
  est <- coef(sa_att)[1]
  se_val <- se(sa_att)[1]
  
  tibble(
    outcome = y,
    method = "Sun-Abraham",
    estimate = est,
    se = se_val,
    conf.low = est - 1.96 * se_val,
    conf.high = est + 1.96 * se_val,
    p.value = 2 * pnorm(abs(est / se_val), lower.tail = FALSE)
  )
}

# ------------------------------------------------------------
# Helper: Callaway-Sant'Anna
# ------------------------------------------------------------

make_cs_data <- function(y, data = df) {
  data %>%
    mutate(
      id_num = as.numeric(as.integer(factor(census_geoid))),
      year = as.numeric(year),
      g = if_else(ever_treated == 1, as.numeric(first_entry_year), 0),
      weight = as.numeric(weight),
      outcome_y = as.numeric(.data[[y]])
    ) %>%
    select(id_num, year, g, weight, outcome_y) %>%
    filter(
      !is.na(id_num),
      !is.na(year),
      !is.na(g),
      !is.na(weight),
      !is.na(outcome_y),
      weight > 0
    ) %>%
    as.data.frame()
}

run_cs <- function(y, data = df) {
  df_cs <- make_cs_data(y, data)
  
  att <- did::att_gt(
    yname = "outcome_y",
    tname = "year",
    idname = "id_num",
    gname = "g",
    data = df_cs,
    panel = TRUE,
    allow_unbalanced_panel = TRUE,
    control_group = "nevertreated",
    xformla = ~ 1,
    est_method = "reg",
    weightsname = "weight",
    bstrap = FALSE,
    faster_mode = FALSE
  )
  
  simple <- did::aggte(att, type = "simple", na.rm = TRUE)
  dynamic <- did::aggte(att, type = "dynamic", na.rm = TRUE)
  
  list(att_gt = att, simple = simple, dynamic = dynamic)
}

extract_cs <- function(cs_obj, y) {
  est <- cs_obj$simple$overall.att
  se_val <- cs_obj$simple$overall.se
  
  tibble(
    outcome = y,
    method = "Callaway-Sant'Anna",
    estimate = est,
    se = se_val,
    conf.low = est - 1.96 * se_val,
    conf.high = est + 1.96 * se_val,
    p.value = 2 * pnorm(abs(est / se_val), lower.tail = FALSE)
  )
}

# ------------------------------------------------------------
# Run all models
# ------------------------------------------------------------

twfe_models <- map(outcomes, run_twfe)
names(twfe_models) <- outcomes

sa_models <- map(outcomes, run_sa)
names(sa_models) <- outcomes

cs_models <- map(outcomes, run_cs)
names(cs_models) <- outcomes

# ------------------------------------------------------------
# Extract comparison table
# ------------------------------------------------------------

twfe_results <- map2_dfr(twfe_models, names(twfe_models), extract_twfe)
sa_results <- map2_dfr(sa_models, names(sa_models), extract_sa)
cs_results <- map2_dfr(cs_models, names(cs_models), extract_cs)

did_results_all <- bind_rows(
  twfe_results,
  sa_results,
  cs_results
) %>%
  mutate(
    outcome_label = recode(outcome, !!!outcome_labels),
    method = factor(
      method,
      levels = c("TWFE", "Sun-Abraham", "Callaway-Sant'Anna")
    ),
    outcome_label = factor(
      outcome_label,
      levels = outcome_labels[outcomes]
    )
  )

did_results_all %>%
  arrange(outcome_label, method) %>%
  print(n = Inf)
```

```{r multi-outcome-did-plot, message=FALSE, warning=FALSE}
ggplot(did_results_all, aes(x = method, y = estimate)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  geom_point(size = 2.5) +
  geom_errorbar(
    aes(ymin = conf.low, ymax = conf.high),
    width = 0.15,
    linewidth = 0.7
  ) +
  coord_flip() +
  facet_wrap(~ outcome_label, scales = "free_x") +
  labs(
    title = "Estimated Effects of Data Center Entry Across DiD Estimators",
    subtitle = "TWFE, Sun-Abraham, and Callaway-Sant'Anna",
    x = NULL,
    y = "Estimated ATT ($/MWh)"
  ) +
  theme_minimal(base_size = 12)
```

```{r save-multi-outcome-results, message=FALSE, warning=FALSE}
dir.create("output", showWarnings = FALSE)

write_csv(
  did_results_all,
  "output/did_results_all_outcomes.csv"
)
```